What Can Stack Traces Tell us about Exceptions in Java?

### 2014-07-07 android-stacks.tex
mprovements on the Introduction

Bug issues have called the attention of research community and have been the target of several 
researches [][][]. A subtype of a bug issue are the ones related to system 
crashes, also called crash reports such reports can be automatically generated 
as the system crashes or can be manually reported by developers/users [complete]...

One of the most important elements of a crash report genereted eigther manually
or automatically is the failure stack trace [ref-paper], which tipically consists 
of an ordered list of methods that were active on the call stack before the failure occurred.

Such failure stack traces are a useful source of 
information about system crashes, and are often used to
support developers in debugging~\cite{schroter2010stack}. Moreover, information
from stack traces combined with crash and bug reports can be used to support bug
classification and clustering~\cite{wang2013improving, kim2011crash,
dhaliwal2011classifying}, fault-proneness prediction
models~\cite{kim2013predicting} and even automated bug
fixing~\cite{sinha2009fault} tools.

The exception handling mechanism~\cite{goodenough1975exception},
 is one of the most used schemes for detecting and recovering from such 
failures. Exceptions may arrise as a consequence of faults in the application
itself (e.g., access of null references), noisy user inputs or faults in
underlying middleware or hardware.

Although moder application have to deal with a similar and increasing number of 
source of failures, existing research has focused on the analysis of a 
single system at a time.

Although the exception handling mechanism has been embedded in several 
programming languages (e.g. Java, C++, C\#) and was the target of several studies
(e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting, garcia2001comparative,cabral2007exception,coelho2011unveiling}),
the exception handling code is often generally poorly understood and the least
tested part of software systems.

Specially in the open-source environment, where faults are usualy not hidden in a private trackers,  it would be intresting if developers could learn from other projects mistakes. Moreover, interesting knowledge may emerge from a cross-project crash analysis.

Hence, in such context where we have access to wealth of information available
on open-source repositories such as GitHub, some questions arise: What are the common
characteristics of Github crash reports?

. The ExceptionMiner tool is a contribution in this direction (to enable a cross-project defect analysis).

In Java, when the program fails due to an uncaught exception, it automatically
terminates, while the system prints a stack trace to the console.

A stack trace
includes an explanatory message and the execution stack frame when the exception
occurred. A typical Java stack trace consists of an ordered list of methods that
were active on the call stack before the exception occurred. Stack traces
are a useful source of information about system crashes, and are often used to
support developers in debugging~\cite{schroter2010stack}. Moreover, information
from stack traces combined with crash and bug reports can be used to support bug
classification and clustering~\cite{wang2013improving, kim2011crash,
dhaliwal2011classifying}, fault-proneness prediction
models~\cite{kim2013predicting} and even automated bug
fixing~\cite{sinha2009fault} tools.

@@

@@

The main disadvantage of return codes is the fact that the programmer can just
ignore the return code and let the program execute in inconsistent state. The
exception mechanism~\cite{goodenough1975exception} was proposed as a way of
preventing errors from being ignored and was embedded in many mainstream
programming languages ~\cite{garcia2001comparative} since then.

@@

@@

As mentioned before, this study amined at mining information available on GitHub issues,
more specifically, extracting and distilling the stack traces embedded on GItHub issues.

Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
on which repository it was created.

Therefore, this study is based on the assumption if an stack trace information embedded
 in an issue, it contain relevant information concerning the exception use -  regardless the
 lables associated to the issue.

@@

This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
offered through the Github API.  GHTorrent has been collecting data from Git since 
February 2012. Up to Dec 2013, and when we queried GHTorrent, it included 332,864 non-fork Java
repositories \footnote {A Java project is specified in GitHub with a
specific label.}  from which only 44,323 have contained at least issue report in their lifetime. From this set only 16,837 were watched by external users. To ensure that our work targets projects that are openly used by
users other than their developers, we focused our analysis on just those projects.

From this set of 16.836 Java projects which overall contained 356.057 issues, we retrieved the metadata related to each Java project (e.g., repository's names and short descriptions), and the content of every issue and issue comment and stored in a relational database.

Using the heuristics presented in Section~\ref{sec:android}, we selected 482 repositories featuring Android projects on which we perform manual inspection was performed.

\subsection{Stack Traces on Issues}

@@

Using the ExceptionMiner tool, developed in this work, we could extract the exception stack traces defined on issues and issue comments. Executing ExceptionMiner tool each issue and issue comment defined on the 16,836 projects considered in this study, we could observe that 3,758 projects (22,32\%) contained at least 1 issue on which a stack trace was found.

Overall the number of issues on which a stack trace was found was 21,013 from which 28,800 stack traces were extracted and analyzed - some issues defined more than one stack trace.

We then analyzed the lables associated to such issues, and we discovered that only 5,196 were labled as defect ~\footnote{the dedect lables considered in this work were every lable containing as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7% of the issues on which a stack trace was found), most of the issues on which stack traces were found contained no lable. Such observation supports our assumption to consider not only the issues related with defect issues on this analysis.

\subsection{Android Apps Selection}
\label{sec:android}

@@

This set of 482 projects contained overall 32,582 issues from which 4,208 stacktraces 
were extracted. From this set of issues 44,9% were labled as defect.

@@

\noindent\emph{StackTraceMiner and Distiller} The first step in the process is
mining references to exceptions and stack traces embedded in issues, distilling
the information that composes a stack trace and storing the results in a
relational database. Some of the attributes extracted from the stack traces were:
 the exception signaler, the root cause, and the root cause wrappers. 
The tool is based on a combination between a regular expression based parser 
with heuristics able to identify and filter exception names and stack traces inline with text. In
contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work can extract all causes related to an exception on a stack,
and stack traces embedded in logs files.

@@

%\subsection{Phase 1: Analyzing Pure Stack Trace Information of Java Programs}

\section{RQ1: What is the prevalence of stack traces on reported issues?}

\section{RQ2: What are the common characteristics of such stack traces?}

@@

\subsection{Top 10 most common exceptions in Java Projects}

Table~\ref{tab:toptenjava} illustrates the top 10 most common root causes which
represents more than 50\% of the analyzed set. The table illustrates  i) the
number of times each exception appeared as the root cause reported on the exception stack trace
(\#occurrences), ii) the number of distinct projects reporting such causes on
stack traces, and iii) the exception popularity among projects.

\begin{table}
\begin{tabular}{lrrrr}
    \hline
    \bfseries{Exception} &  \multicolumn{2}{c}{\bfseries{Occurrences}} &  \multicolumn{2}{c}{\bfseries{Projects}}\\
    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } \\
    \hline
java.lang.NullPointerException             & 7578 & 26,31\% & 1671 & 44,47 \\
java.lang.IllegalArgumentException         & 1573 & 5,46\%  & 670  & 17,83\% \\
java.lang.ClassNotFoundException           & 1203 & 4,18\%  & 514  & 13,68\% \\
java.lang.RuntimeException                 & 1093 & 3,80\%  & 411  & 10,94\% \\
java.lang.IllegalStateException            & 964  & 3,35\%  & 431  & 11,47\% \\
java.lang.ClassCastException               & 886  & 3,08\%  & 408  & 10,86\% \\
java.lang.ArrayIndexOutOfBoundsException   & 834  & 2,90\%  & 345  & 9,18\% \\
java.lang.NoSuchMethodError                & 822  & 2,85\%  & 369  & 9,82\% \\
java.io.IOException                        & 656  & 2,28\%  & 337  & 8,97\% \\
java.lang.OutOfMemoryError                 & 601  & 2,09\%  & 235  & 6,25\% \\
%java.lang.NoClassDefFoundError            & 554  & 1,92\%  & 299  & 7,96\% \\
%java.util.ConcurrentModificationException & 465  & 1,61\%  & 180  & 4,79\% \\
%java.lang.IndexOutOfBoundsException       & 445  & 1,55\%  & 205  & 5,46\% \\
\hline
All	 &	23956 &	100\%	& 3758 & 100\%	 \\
\hline
  \end{tabular}
\caption{Top 10 most popular exceptions on issues in Java projects}
\label{tab:toptenjava}
\end{table}

We can observe that the NullPointerException was the one with the higher number
of occurrences (26,3\% of all stacks were caused by NullPointerExceptions) and
the one which had the highest popularity across projects. Actually, most of the
exceptions in the top 10 list (except java.lang.IllegalArgumentException,
java.io.IOException and java.lang.RuntimeException) are exceptions implicitly
thrown by Java Virtual Machine, when an expression violates the normal
semantics of the Java programming language (e.g. ArrayIndexOutOfBounds) or due
to an internal error or resource limitation (OutOfMemoryError).

The high prevalence of NullPointerExceptions, and the other implicitly-thrown
exceptions  is aligned with the findings other works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs related to exception handling were caused by NullPointerException; other works on robustness testing~\cite{maji2012empirical,csallner2004jcrasher}
showed that most of the automatically
detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
environment (as the ones found in this study).

\subsection{Top 10 most common exceptions in Android Projects}

@@

\subsection{Classification of the main causes of top exceptions}

To better understand the main causes of such exceptions, we evaluated semantics
related to the top 100 root causes reported on stack-issues of both Java dataset
and Android subset (which correspond to approximately 77\% of all exceptions
reported reported on Java stack-issues and 95\% of all exceptions reported on
Android stack-issues), and classified them according to the the categories
presented in Table~\ref{tab:tophundrend}.

@@

\section{RQ3.  What kinds of exception (mis)-use emerge from stack trace analysis? }

@@

>>>>

Bug issues have called the attention of the research community and have been the target of increasing 
researches [][][]. A specific type of bug issue is the one related to system 
crashes.  Such issues usually contain an exception stack trace [ref-paper], which tipically consists 
of an ordered list of methods that were active on the call stack before the failure occurred [ref-bug-issues].

Exception stack traces are a useful source of 
information about system crashes, and are often used to
support developers in debugging~\cite{schroter2010stack}. Moreover, information
from stack traces combined with crash and bug reports can be used to support bug
classification and clustering~\cite{wang2013improving, kim2011crash,
dhaliwal2011classifying}, fault-proneness prediction
models~\cite{kim2013predicting} and even automated bug
fixing~\cite{sinha2009fault} tools.

Although modern applications have to deal with increasinng and similar 
sources of failures - that arise as a consequence of faults in the application
itself (e.g., access of null references), noisy user inputs or faults in
underlying middleware or hardware -  on bug classificationthe current research on 
exploring bug issue information are structured on  a system at a time basis.

Specially  in a context where we have access to wealth of information available
on open-source repositories such as GitHub, interesting knowledge may emerge 
from a cross-project crash analysis. Hence such initial questions arise: What are the common
characteristics of the stack traces reported on GitHub? Which patterns of exception (mis)use may emerge?

@@

  \item  It performs a large scale analysis of Java stack traces and how
    they can reveal bad-practices on the use of exceptions in Java.

@@

The main disadvantage of return codes is the fact that the programmer can just
ignore the return code and let the program execute in inconsistent state.

The exception mechanism~\cite{goodenough1975exception} was proposed as a way of
preventing errors from being ignored and was embedded in many mainstream
programming languages ~\cite{garcia2001comparative} since then.

@@

.......

Although exception handling mechanism is the most used technique to deal with such 
failures - been embedded in several programming languages (e.g. Java, C++, C\#) -  
the exception handling code is often generally poorly understood and the least
tested part of software systems.This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes
(e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting, garcia2001comparative,cabral2007exception,coelho2011unveiling}).

........

@@

As mentioned before, this study amined at mining information available on GitHub issues,
more specifically, extracting and distilling the stack traces embedded on GItHub issues. 
Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
on which repository it was created. Therefore, this study is based on the assumption 
if an stack trace information embedded in an issue, it contain relevant information
 concerning the exception use -  regardless the lables associated to the issue.

@@

\subsubsection{GHTorrent Dataset} This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
offered through the Github API.  GHTorrent has been collecting data from Git since 
February 2012. Up to Dec 2013, and when we queried GHTorrent, it included 332,864 non-fork Java
repositories \footnote {A Java project is specified in GitHub with a
specific label.}  from which only 44,323 have contained at least issue report in their lifetime. From this set only 16,837 were watched by external users. To ensure that our work targets projects that are openly used by
users other than their developers, we focused our analysis on just those projects.

From this set of 16.836 Java projects which overall contained 356.057 issues, we retrieved the metadata related to each Java project (e.g., repository's names and short descriptions), and the content of every issue and issue comment and stored in a relational database. Using the heuristics presented in Section~\ref{sec:android}, we selected 482 repositories featuring Android projects on which we perform manual inspection was performed.

\subsubsection{Stack Traces on Issues}

@@

Using the ExceptionMiner tool, developed in this work, we could extract the exception stack traces defined on issues and issue comments. Executing ExceptionMiner tool each issue and issue comment defined on the 16,836 projects considered in this study, we could observe that 3,758 projects (22,32\%) contained at least 1 issue on which a stack trace was found. Overall the number of issues on which a stack trace was found was 21,013 from which 28,800 stack traces were extracted and analyzed - some issues defined more than one stack trace.

We then analyzed the lables associated to such issues, and we discovered that only 5,196 were labled as defect ~\footnote{the dedect lables considered in this work were every lable containing as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7\% of the issues on which a stack trace was found), most of the issues on which stack traces were found contained no lable. Such observation supports our assumption to consider not only the issues related with defect issues on this analysis.

\subsubsection{Android Apps Selection}
\label{sec:android}

@@

This set of 482 projects contained overall 32,582 issues from which 4,208 stacktraces 
were extracted. From this set of issues 44,9\% were labled as defect.

@@

\noindent\emph{StackTraceMiner and Distiller} The first step in the process is
mining references to exceptions and stack traces embedded in issues, distilling
the information that composes a stack trace and storing the results in a
relational database. Some of the information extracted from the stack traces were:
 the exception being thrown, its signaler, the exception root cause, its corresponding signaler,
and packages related to each one of these fields.
The tool is based on a combination between a regular expression based parser 
with heuristics able to identify and filter exception names and stack traces inline with text. In
contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work can extract all causes related to an exception on a stack,
and stack traces embedded in logs files.

@@

\section{RQ1: What are the common characteristics of such stack traces?}

@@

\subsection{Characterizing the Most Common Exceptions}

After distilling the information available on the 28,800 stack traces were extracted from the 
issues and issues comments of 16,836 Java projects: 607 distinct root causes were found, 
and  2,026 distinct top level exceptions were found.

From this set the java.lang.NullPointerException  was the root cause most 
reported;  7578 stack traces had a java.lang.NullPointerException  as root cause,
 which corresponds of 26,31\%  of all analyzed stack traces.

If we perform a cross-project analysis, the java.lang.NullPointerException, we could observe that such 
exception was reported on 1,671 projects (which represents 44,47% of all projects reporting at least one stack trace).

The high prevalence of NullPointerExceptions is aligned with the findings other works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs related to exception handling were caused by NullPointerException; other works on robustness testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
environment (as the ones found in this study).

To better understand the root causes reported on stacks, we performed a categorization 
study on which the top 100 root causes reported on stack-issues were manually inspected.
The characterization process consisted of classifiying each exception in one of the
chategories listed on Table X, based on the inspection of source code, Javadoc, 
or additional documentation related to the exception.

\begin{table}
  \centering
  \begin{tabular}{|p{2cm}| p{5cm}|}
    \hline
    \bfseries{Category} & \bfseries{Description} \\
    \hline
      Programming logic &  Exceptions implicitly  thrown by the JVM when the 
evaluation of an expression violates the normal semantics of language (e.g., 
out-of-bounds array index, division-by-zero, access to a null reference). 
Such exceptions are usually defined on java.lang and java.util packages. \\ \hline
      Resources (IO)                         & Exception releated to the manipulation of IO resources or device  (e.g., file, network) \\ \hline
      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
      Reflection                             & Exceptions related to Java reflection library  \\ \hline
      Specific                & Exceptions created for specific applications, frameworks or libraries \\ \hline
      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
    \hline
  \end{tabular}
  \caption{Categories description.}
  \label{tab:categories}
\end{table}

The top 100 root causes correspond to approximately 77\% of all stack traces
reported reported on Java stack-issues and 95\% of all exceptions reported on
Android stack-issues. Table~\ref{tab:tophundrend} illustrates the result of such 
classification.

\subsection{Investigating Android Subset}

@@

@@

\section{RQ2.  What kinds of exception (mis)-use emerge from stack trace analysis? }

@@

\section {Old Introduction}

In Java, when the program fails due to an uncaught exception, it automatically
terminates, while the system prints a stack trace to the console. A stack trace
includes an explanatory message and the execution stack frame when the exception
occurred. A typical Java stack trace consists of an ordered list of methods that
were active on the call stack before the exception occurred. Stack traces
are a useful source of information about system crashes, and are often used to
support developers in debugging~\cite{schroter2010stack}. Moreover, information
from stack traces combined with crash and bug reports can be used to support bug
classification and clustering~\cite{wang2013improving, kim2011crash,
dhaliwal2011classifying}, fault-proneness prediction
models~\cite{kim2013predicting} and even automated bug
fixing~\cite{sinha2009fault} tools.
